% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
\documentclass[
  english,
  10pt,
  a4paper,
  oneside,
  final]{article}
\usepackage{xcolor}
\usepackage[left=2.245cm,right=2.245cm,top=2.5cm,bottom=2.5cm]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{5}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{246,248,250}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{#1}
\newcommand{\AttributeTok}[1]{#1}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.62,0.41,0.03}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.62,0.41,0.03}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.62,0.41,0.03}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{#1}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.38,0.49}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.00,0.38,0.49}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.62,0.41,0.03}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.62,0.41,0.03}{#1}}
\newcommand{\FunctionTok}[1]{#1}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.38,0.49}{#1}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.48,0.24,0.57}{#1}}
\newcommand{\OtherTok}[1]{#1}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.44,0.31,0.22}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.62,0.41,0.03}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.62,0.41,0.03}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.62,0.41,0.03}{#1}}
\newcommand{\VariableTok}[1]{#1}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.62,0.41,0.03}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.42,0.45,0.49}{\textbf{#1}}}
\usepackage{longtable,booktabs,array}
\newcounter{none} % for unnumbered tables
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\ifLuaTeX
  \usepackage{luacolor}
  \usepackage[soul]{lua-ul}
\else
  \usepackage{soul}
\fi
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic,shorthands=off]{babel}
\else
\usepackage[bidi=default,shorthands=off]{babel}
\fi
\ifLuaTeX
  \usepackage{selnolig} % disable illegal ligatures
\fi
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage[normalem]{ulem}
\providecommand{\CSLIndent}{}
\renewcommand{\CSLIndent}[1]{\hfill\break\hspace*{\cslhangindent}\ignorespaces#1}

% adjust quote indent
\renewenvironment{quote}{\list{}{\leftmargin=0.2in}\item[]}{\endlist}

% https://github.com/cplusplus/draft/blob/97b615a5a6ab0598b624ee05402c531d0421cff6/source/styles.tex#L127-L133
\renewcommand{\labelitemi}{---}
\renewcommand{\labelitemii}{---}
\renewcommand{\labelitemiii}{---}
\renewcommand{\labelitemiv}{---}

% https://github.com/cplusplus/draft/blob/97b615a5a6ab0598b624ee05402c531d0421cff6/source/layout.tex#L32-L68
\makeatletter
\newcommand{\pnum}[1]{%
\noindent\makebox[0pt][l]{\makebox[0pt][r]{%
\scriptsize\raisebox{.7ex}{#1}%
\hspace{\@totalleftmargin}\quad%
}}\ignorespaces}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Deduced return type for overriding functions via `auto`},
  pdflang={en},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={blue},
  urlcolor={blue},
  pdfcreator={LaTeX via pandoc}}
\hypersetup{pdfcreator={mpark/wg21}}

\makeatletter
\def\@maketitle{
  \newpage \null \vskip 2em
  {\center \LARGE \@title \par}
  \vskip 1.5em
  \begin{flushright}
    \begin{tabular}{ll}
Document \#:&DnnnnR0\\
Date:       &\@date\\
Project:    &Programming Language C++\\
Audience:   &SG17 (EWG Incubator)\\&EWG (Evolution Working Group)\\
Reply-to:   \@author
    \end{tabular}
  \end{flushright}
}
\makeatother

\title{Deduced return type for overriding functions via
\VERB|\KeywordTok{auto}|}
\author{&Alex
Tsvetanov\\&<\href{mailto:alex_tsvetanov_2002@abv.bg}{\nolinkurl{alex\_tsvetanov\_2002@abv.bg}}>\\}
\date{2026-06-25}

\begin{document}
\hypertarget{title-block-header}{\label{title-block-header}}
\maketitle

{
\hypersetup{linkcolor=black}
\setcounter{tocdepth}{2}
\hypertarget{toctitle}{\label{toctitle}}
\tableofcontents
}
\section{\texorpdfstring{Abstract\hyperref[abstract]{}}{Abstract}}\label{abstract}

A function that overrides a virtual function must, today, repeat the
overridden function's return type --- either by spelling it out or by
reconstructing it with a trailing-return-type such as
\VERB|\OperatorTok{{-}\textgreater{}} \KeywordTok{decltype}\OperatorTok{(}\NormalTok{std}\OperatorTok{::}\NormalTok{declval}\OperatorTok{\textless{}}\NormalTok{Base}\OperatorTok{\textgreater{}().}\NormalTok{f}\OperatorTok{())}|.
This paper proposes that a function declared with the bare placeholder
\VERB|\KeywordTok{auto}| \emph{and} the \VERB|\KeywordTok{override}|
specifier deduce its return type to be \textbf{exactly the return type
of the function it overrides}. Because \VERB|\KeywordTok{override}|
already guarantees that a matching base function exists, the deduction
is unambiguous, requires no function body, and the resulting function is
an ordinary virtual function in every other respect.

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\centering\arraybackslash\textbf{Before}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\centering\arraybackslash\textbf{After}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\begin{minipage}[t]{\linewidth}\raggedright
\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{struct}\NormalTok{ A }\OperatorTok{\{}
  \KeywordTok{virtual} \DataTypeTok{int}\NormalTok{ test}\OperatorTok{()} \OperatorTok{\{} \ControlFlowTok{return} \DecValTok{5}\OperatorTok{;} \OperatorTok{\}}
\OperatorTok{\};}

\KeywordTok{struct}\NormalTok{ B }\OperatorTok{:}\NormalTok{ A }\OperatorTok{\{}
  \CommentTok{// Return type must be restated, either literally...}
  \DataTypeTok{int}\NormalTok{ test}\OperatorTok{()} \KeywordTok{override} \OperatorTok{\{} \ControlFlowTok{return} \DecValTok{6}\OperatorTok{;} \OperatorTok{\}}

  \CommentTok{// ...or reconstructed when A::test\textquotesingle{}s type is awkward to name:}
  \KeywordTok{auto}\NormalTok{ test2}\OperatorTok{()} \OperatorTok{{-}\textgreater{}} \KeywordTok{decltype}\OperatorTok{(}\NormalTok{std}\OperatorTok{::}\NormalTok{declval}\OperatorTok{\textless{}}\NormalTok{A}\OperatorTok{\textgreater{}().}\NormalTok{test2}\OperatorTok{())}
      \KeywordTok{override} \OperatorTok{\{} \ControlFlowTok{return} \DecValTok{6}\OperatorTok{;} \OperatorTok{\}}
\OperatorTok{\};}
\end{Highlighting}
\end{Shaded}
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{struct}\NormalTok{ A }\OperatorTok{\{}
  \KeywordTok{virtual} \DataTypeTok{int}\NormalTok{ test}\OperatorTok{()} \OperatorTok{\{} \ControlFlowTok{return} \DecValTok{5}\OperatorTok{;} \OperatorTok{\}}
\OperatorTok{\};}

\KeywordTok{struct}\NormalTok{ B }\OperatorTok{:}\NormalTok{ A }\OperatorTok{\{}
  \CommentTok{// Return type is taken from the overridden function.}
  \KeywordTok{auto}\NormalTok{ test}\OperatorTok{()} \KeywordTok{override} \OperatorTok{\{} \ControlFlowTok{return} \DecValTok{6}\OperatorTok{;} \OperatorTok{\}}

  \KeywordTok{auto}\NormalTok{ test2}\OperatorTok{()} \KeywordTok{override} \OperatorTok{\{} \ControlFlowTok{return} \DecValTok{6}\OperatorTok{;} \OperatorTok{\}}
\OperatorTok{\};}
\end{Highlighting}
\end{Shaded}
\end{minipage} \\
\end{longtable}
}

\section{\texorpdfstring{Introduction\hyperref[introduction]{}}{Introduction}}\label{introduction}

C++ requires an overriding virtual function to repeat the return type of
the function it overrides, even though that type is fixed by the
language and carries no new information. This paper proposes a small,
self-contained extension: a function declared with the bare placeholder
\VERB|\KeywordTok{auto}| together with the \VERB|\KeywordTok{override}|
specifier takes its return type directly from the function it overrides.
The change is a pure language extension --- it makes a presently
ill-formed spelling well-formed, alters no existing program, and
requires no library, ABI, or object-model changes. The remainder of the
paper motivates the feature, discusses the design, and provides proposed
wording.

\section{\texorpdfstring{Revision
history\hyperref[revision-history]{}}{Revision history}}\label{revision-history}

\subsection*{\texorpdfstring{R0\hyperref[r0]{}}{R0}}\label{r0}
\addcontentsline{toc}{subsection}{R0{}}

\begin{itemize}
\tightlist
\item
  Initial revision.
\end{itemize}

\section{\texorpdfstring{Motivation and
scope\hyperref[motivation-and-scope]{}}{Motivation and scope}}\label{motivation-and-scope}

\subsection{\texorpdfstring{The
problem\hyperref[the-problem]{}}{The problem}}\label{the-problem}

When overriding a virtual function, the override's return type is
constrained by the language: it must be \emph{identical} to the base
function's return type, or a legal covariant type
(\href{https://eel.is/c++draft/class.virtual}{{[}class.virtual{]}}). The
programmer therefore gains nothing by restating the return type --- it
carries no new information, it cannot legally differ (except for
covariance), and it must be kept in sync by hand.

This redundancy becomes actively harmful when the base return type is
verbose, dependent, or obtained through metaprogramming. The idiomatic
way to ``inherit'' the return type today is:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{struct}\NormalTok{ B }\OperatorTok{:}\NormalTok{ A }\OperatorTok{\{}
  \KeywordTok{auto}\NormalTok{ test}\OperatorTok{()} \OperatorTok{{-}\textgreater{}} \KeywordTok{decltype}\OperatorTok{(}\NormalTok{std}\OperatorTok{::}\NormalTok{declval}\OperatorTok{\textless{}}\NormalTok{A}\OperatorTok{\textgreater{}().}\NormalTok{test}\OperatorTok{())} \KeywordTok{override} \OperatorTok{\{} \ControlFlowTok{return} \DecValTok{6}\OperatorTok{;} \OperatorTok{\}}
\OperatorTok{\};}
\end{Highlighting}
\end{Shaded}

That trailing-return-type is pure boilerplate: it names the base class
explicitly (\VERB|\NormalTok{A}|), it duplicates the function name
(\VERB|\NormalTok{test}|), and it is fragile under refactoring ---
renaming the base, or moving the function through the hierarchy,
silently requires the expression to be updated. Worse,
\VERB|\NormalTok{std}\OperatorTok{::}\NormalTok{declval}|/\VERB|\KeywordTok{decltype}|
is an advanced idiom that obscures a trivial intent: \emph{``return
whatever the base returns.''}

\subsection{\texorpdfstring{What this paper
proposes\hyperref[what-this-paper-proposes]{}}{What this paper proposes}}\label{what-this-paper-proposes}

Allow a virtual override to be written with the bare placeholder type
\VERB|\KeywordTok{auto}| and no trailing-return-type:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{struct}\NormalTok{ B }\OperatorTok{:}\NormalTok{ A }\OperatorTok{\{}
  \KeywordTok{auto}\NormalTok{ test}\OperatorTok{()} \KeywordTok{override} \OperatorTok{\{} \ControlFlowTok{return} \DecValTok{6}\OperatorTok{;} \OperatorTok{\}}
\OperatorTok{\};}
\end{Highlighting}
\end{Shaded}

When a function is declared this way, its return type is determined to
be \textbf{the return type of the unique function it overrides} in a
direct or indirect base class. The determination happens at the point of
declaration from the override relationship --- it does \textbf{not}
depend on the function body --- so the function behaves exactly as if
the return type had been written out by hand.

\subsection{\texorpdfstring{Why this is safe and
small\hyperref[why-this-is-safe-and-small]{}}{Why this is safe and small}}\label{why-this-is-safe-and-small}

The proposal deliberately ties the feature to the
\VERB|\KeywordTok{override}| specifier (see {[}design.trigger{]}). The
\VERB|\KeywordTok{override}| keyword already makes the program
ill-formed unless the function overrides a base virtual function
(\href{https://eel.is/c++draft/class.virtual\#4}{{[}class.virtual{]}/4}),
so by construction there is always exactly one return type to copy. The
feature therefore introduces no new failure mode that
\VERB|\KeywordTok{override}| did not already diagnose: a function marked
\VERB|\KeywordTok{override}| that fails to override anything is
ill-formed today, and remains ill-formed under this proposal.

\section{\texorpdfstring{Design
discussion\hyperref[design-discussion]{}}{Design discussion}}\label{design-discussion}

\subsection{\texorpdfstring{The feature is triggered only by
\texttt{override}\hyperref[design.trigger]{}}{The feature is triggered only by override}}\label{design.trigger}

Deduction applies only to a function declared with \textbf{both} the
placeholder type \VERB|\KeywordTok{auto}| \textbf{and} the
\VERB|\KeywordTok{override}| specifier. We do \emph{not} extend it to
any function that merely happens to match a base virtual signature.

Tying the feature to \VERB|\KeywordTok{override}| has three benefits:

\begin{itemize}
\tightlist
\item
  \textbf{A return type always exists to copy.}
  \VERB|\KeywordTok{override}| guarantees a matching base function, so
  the deduction can never silently fail to find a source.
\item
  \textbf{Intent is explicit.} The reader sees
  \VERB|\KeywordTok{override}| and knows the function is bound to a base
  contract; the return type following from that contract is
  unsurprising.
\item
  \textbf{No accidental coupling.} A non-\VERB|\KeywordTok{override}|
  function with a deduced return type continues to deduce from its body,
  exactly as today. Adding or removing a base virtual function never
  silently changes the meaning of an unrelated \VERB|\KeywordTok{auto}|
  function.
\end{itemize}

\subsection{\texorpdfstring{The deduced type is the overridden
function's type,
exactly\hyperref[design.exact]{}}{The deduced type is the overridden function's type, exactly}}\label{design.exact}

The return type is copied verbatim from the overridden function. It is
\textbf{not} deduced from the override's \VERB|\ControlFlowTok{return}|
statements. Two consequences:

\begin{itemize}
\item
  The function body is irrelevant to the return type. A declaration with
  no definition (e.g.~an out-of-line definition) is fully supported:

  \begin{Shaded}
  \begin{Highlighting}[]
  \KeywordTok{struct}\NormalTok{ B }\OperatorTok{:}\NormalTok{ A }\OperatorTok{\{}
    \KeywordTok{auto}\NormalTok{ test}\OperatorTok{()} \KeywordTok{override}\OperatorTok{;}     \CommentTok{// return type is A::test\textquotesingle{}s type: int}
  \OperatorTok{\};}
  \DataTypeTok{int}\NormalTok{ B}\OperatorTok{::}\NormalTok{test}\OperatorTok{()} \OperatorTok{\{} \ControlFlowTok{return} \DecValTok{6}\OperatorTok{;} \OperatorTok{\}} \CommentTok{// out{-}of{-}line definition; type already fixed}
  \end{Highlighting}
  \end{Shaded}
\item
  \textbf{Covariance is not synthesized.} Today an override of
  \VERB|\NormalTok{Base}\OperatorTok{*}\NormalTok{ f}\OperatorTok{()}|
  may return \VERB|\NormalTok{Derived}\OperatorTok{*}|. Under this
  proposal,
  \VERB|\KeywordTok{auto}\NormalTok{ f}\OperatorTok{()} \KeywordTok{override}|
  yields exactly \VERB|\NormalTok{Base}\OperatorTok{*}|. A programmer
  who wants a covariant return type must still write it explicitly:

  \begin{Shaded}
  \begin{Highlighting}[]
  \KeywordTok{struct}\NormalTok{ A }\OperatorTok{\{} \KeywordTok{virtual}\NormalTok{ A}\OperatorTok{*}\NormalTok{ clone}\OperatorTok{();} \OperatorTok{\};}
  \KeywordTok{struct}\NormalTok{ B }\OperatorTok{:}\NormalTok{ A }\OperatorTok{\{}
    \KeywordTok{auto}\NormalTok{  clone}\OperatorTok{()} \KeywordTok{override}\OperatorTok{;}   \CommentTok{// returns A*  (base type, exactly)}
  \NormalTok{  B}\OperatorTok{*}\NormalTok{    clone2}\OperatorTok{()} \KeywordTok{override}\OperatorTok{;}  \CommentTok{// covariant return type, written out (unchanged today)}
  \OperatorTok{\};}
  \end{Highlighting}
  \end{Shaded}
\end{itemize}

This keeps the rule trivially simple --- \emph{the override's type is
the overridden function's type} --- and preserves covariance as an
explicit, opt-in act rather than something the compiler infers.

\subsection{\texorpdfstring{Ambiguity across multiple base
functions\hyperref[ambiguity-across-multiple-base-functions]{}}{Ambiguity across multiple base functions}}\label{ambiguity-across-multiple-base-functions}

A function may override more than one base function under multiple
inheritance. If the overridden functions do not all have the same return
type, there is no single type to copy and the program is ill-formed:

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{struct}\NormalTok{ A }\OperatorTok{\{} \KeywordTok{virtual} \DataTypeTok{int}\NormalTok{  f}\OperatorTok{();} \OperatorTok{\};}
\KeywordTok{struct}\NormalTok{ C }\OperatorTok{\{} \KeywordTok{virtual} \DataTypeTok{long}\NormalTok{ f}\OperatorTok{();} \OperatorTok{\};}
\KeywordTok{struct}\NormalTok{ D }\OperatorTok{:}\NormalTok{ A}\OperatorTok{,}\NormalTok{ C }\OperatorTok{\{}
  \KeywordTok{auto}\NormalTok{ f}\OperatorTok{()} \KeywordTok{override}\OperatorTok{;}  \CommentTok{// ill{-}formed: A::f returns int, C::f returns long}
\OperatorTok{\};}
\end{Highlighting}
\end{Shaded}

Note this case is already ill-formed today for a different reason (an
override cannot simultaneously satisfy two incompatible base return
types), so the proposal does not introduce a new diagnosable situation
here; it only specifies which diagnostic applies.

\subsection{\texorpdfstring{\texttt{decltype(auto)} and other
placeholders\hyperref[decltypeauto-and-other-placeholders]{}}{decltype(auto) and other placeholders}}\label{decltypeauto-and-other-placeholders}

This paper addresses only the bare placeholder \VERB|\KeywordTok{auto}|.
A virtual function declared with
\VERB|\KeywordTok{decltype}\OperatorTok{(}\KeywordTok{auto}\OperatorTok{)}|
or a constrained placeholder
(e.g.~\VERB|\NormalTok{std}\OperatorTok{::}\NormalTok{integral }\KeywordTok{auto}|)
remains ill-formed, as today. Extending the feature to those spellings
is possible future work (see {[}future{]}) but is intentionally out of
scope to keep the rule minimal.

\subsection{\texorpdfstring{Relationship to the existing prohibition on
virtual
\texttt{auto}\hyperref[relationship-to-the-existing-prohibition-on-virtual-auto]{}}{Relationship to the existing prohibition on virtual auto}}\label{relationship-to-the-existing-prohibition-on-virtual-auto}

Today, a function whose declared return type uses a placeholder type
``shall not be declared \VERB|\KeywordTok{virtual}|'' --- a
deduced-from-body return type is incompatible with the vtable model
because the type must be known to all translation units that see the
class definition. This proposal does \textbf{not} weaken that rationale:
the return type of an
\VERB|\KeywordTok{auto}\NormalTok{ … }\KeywordTok{override}| function is
known from the class definition (it is the base's return type, which is
visible wherever the derived class is complete), so every translation
unit agrees on the type without seeing the body. The proposal therefore
carves a narrow, well-defined exception out of the existing prohibition
rather than removing it.

\subsection{\texorpdfstring{Interaction with \texttt{final}, access,
\texttt{noexcept}, and
qualifiers\hyperref[interaction-with-final-access-noexcept-and-qualifiers]{}}{Interaction with final, access, noexcept, and qualifiers}}\label{interaction-with-final-access-noexcept-and-qualifiers}

\VERB|\KeywordTok{final}|, access specifiers,
\VERB|\KeywordTok{noexcept}|, ref-qualifiers, and cv-qualifiers are
orthogonal to the return type and are unaffected.
\VERB|\KeywordTok{auto}\NormalTok{ f}\OperatorTok{()} \KeywordTok{const} \KeywordTok{override}|,
\VERB|\KeywordTok{auto}\NormalTok{ f}\OperatorTok{()} \OperatorTok{\&\&} \KeywordTok{override}|,
\VERB|\KeywordTok{auto}\NormalTok{ f}\OperatorTok{()} \KeywordTok{noexcept} \KeywordTok{override}|
all behave as expected: the signature determines which base function is
overridden, and that function's return type is copied.

\section{\texorpdfstring{Proposed
wording\hyperref[proposed-wording]{}}{Proposed wording}}\label{proposed-wording}

The proposed changes are relative to the C++ working draft
\href{https://www.open-std.org/jtc1/sc22/wg21/docs/papers/2026/n5046.pdf}{{[}N5046{]}}.
Existing standard text is quoted verbatim;
{{\color[HTML]{006e28}\uline{inserted text is shown like this}}} and
{{\color[HTML]{bf0303}\sout{removed text like this}}}. Paragraph numbers
match N5046 and are to be adjusted editorially. Three paragraphs/tables
are affected.

\subsection{\texorpdfstring{Change 1 --- relax the virtual-function
prohibition\hyperref[change-1-relax-the-virtual-function-prohibition]{}}{Change 1 --- relax the virtual-function prohibition}}\label{change-1-relax-the-virtual-function-prohibition}

In
\textbf{\href{https://eel.is/c++draft/dcl.spec.auto.general}{{[}dcl.spec.auto.general{]}}}
paragraph 17, the placeholder return type is forbidden on \emph{every}
virtual function. The current text reads, in its entirety:

\begin{quote}
\pnum{17} A function declared with a return type that uses a placeholder
type shall not be virtual
(\href{https://eel.is/c++draft/class.virtual}{{[}class.virtual{]}}).
\end{quote}

Modify it to permit the bare-\VERB|\KeywordTok{auto}| +
\VERB|\KeywordTok{override}| form:

\begin{quote}
\pnum{17} A function declared with a return type that uses a placeholder
type shall not be virtual
(\href{https://eel.is/c++draft/class.virtual}{{[}class.virtual{]}}){{\color[HTML]{bf0303}\sout{.}}}
{{\color[HTML]{006e28}\uline{unless the placeholder type is the
\VERB|\NormalTok{auto}| \emph{type-specifier} not followed by a
\emph{trailing-return-type} and the function is declared with the
\VERB|\NormalTok{override}| \emph{virt-specifier}; in that case the
return type of the function is determined from the function it
overrides, as specified in
\href{https://eel.is/c++draft/class.virtual}{{[}class.virtual{]}}.
{[}Note: The return type of such a function is therefore not deduced
from its \VERB|\NormalTok{return}| statements
(\href{https://eel.is/c++draft/dcl.spec.auto.general}{{[}dcl.spec.auto.general{]}}).
--- end note{]}}}}
\end{quote}

\emph{Drafting note.} No other paragraph of
\href{https://eel.is/c++draft/dcl.spec.auto.general}{{[}dcl.spec.auto.general{]}}
needs to change: paragraph 8 (``a program that uses a placeholder type
in a context not explicitly allowed \ldots{} is ill-formed'') already
defers to the per-context rules, and this is the only context that
forbade a virtual function.

\subsection{\texorpdfstring{Change 2 --- specify the return type from
the overridden
function\hyperref[change-2-specify-the-return-type-from-the-overridden-function]{}}{Change 2 --- specify the return type from the overridden function}}\label{change-2-specify-the-return-type-from-the-overridden-function}

In
\textbf{\href{https://eel.is/c++draft/class.virtual}{{[}class.virtual{]}}},
the override-matching rule (paragraph 2) keys on the \emph{name} and
\emph{parameter-type-list} and never on the return type, and paragraph 5
makes \VERB|\KeywordTok{override}| ill-formed unless the function
overrides something. Add a new paragraph immediately after paragraph 8
(the covariant-return-type rule) to determine the return type in the new
case. For reference, paragraph 8 currently begins:

\begin{quote}
\pnum{8} The return type of an overriding function shall be either
identical to the return type of the overridden function or covariant
with the classes of the functions. {[}\ldots{]}
\end{quote}

Insert after it:

{\color[HTML]{006e28}

\begin{quote}
\pnum{(8.*)} If the declared return type of an overriding function
\VERB|\NormalTok{G}| uses the \VERB|\NormalTok{auto}| placeholder type
(\href{https://eel.is/c++draft/dcl.spec.auto}{{[}dcl.spec.auto{]}}),
then \VERB|\NormalTok{G}| is declared with the
\VERB|\NormalTok{override}| \emph{virt-specifier}
(\href{https://eel.is/c++draft/dcl.spec.auto.general}{{[}dcl.spec.auto.general{]}})
and therefore overrides at least one function
(\href{https://eel.is/c++draft/class.virtual\#5}{{[}class.virtual{]}/5}).
If the functions that \VERB|\NormalTok{G}| overrides do not all have the
same return type, the program is ill-formed. Otherwise, the return type
of \VERB|\NormalTok{G}| is that common return type. {[}Note: This return
type is identical to the return type of each overridden function and is
therefore not covariant
(\href{https://eel.is/c++draft/class.virtual\#8}{{[}class.virtual{]}/8})
with it. A covariant return type must be declared explicitly. --- end
note{]}
\end{quote}

}

\emph{Drafting note.} Because paragraph 2's override matching does not
consult the return type, the set of overridden functions --- and hence
``that common return type'' --- is well-defined before
\VERB|\NormalTok{G}|'s return type is known; there is no circularity.

\subsection{\texorpdfstring{Change 3 --- feature-test
macro\hyperref[change-3-feature-test-macro]{}}{Change 3 --- feature-test macro}}\label{change-3-feature-test-macro}

In
\textbf{\href{https://eel.is/c++draft/cpp.predefined}{{[}cpp.predefined{]}}},
add a row to Table {[}tab:cpp.predefined.ft{]} (``Feature-test
macros''), in alphabetical order among the other
\VERB|\NormalTok{\_\_cpp\_}| macros:

{\color[HTML]{006e28}

\begin{quote}
{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.5000}}@{}}
\toprule\noalign{}
\centering\arraybackslash\textbf{Macro name} &
\centering\arraybackslash\textbf{Value} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\VERB|\NormalTok{\_\_cpp\_deduced\_virtual\_override\_return}| &
\VERB|\NormalTok{202606L}| \\
\end{longtable}
}
\end{quote}

}

\emph{Drafting note.} The value \VERB|\DecValTok{202606}\BuiltInTok{L}|
is provisional; the editor assigns the year-and-month
(\VERB|\NormalTok{YYYYMML}|) of the meeting at which the feature is
adopted. The name uses ``virtual'' because the feature applies precisely
to virtual functions: \VERB|\KeywordTok{override}| is permitted only on
a function that overrides a base virtual function
(\href{https://eel.is/c++draft/class.virtual\#5}{{[}class.virtual{]}/5}),
which is therefore itself virtual.

\subsection{\texorpdfstring{Summary of
edits\hyperref[summary-of-edits]{}}{Summary of edits}}\label{summary-of-edits}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\centering\arraybackslash\textbf{\#}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\centering\arraybackslash\textbf{Stable label}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\centering\arraybackslash\textbf{Location}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\centering\arraybackslash\textbf{Edit}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 &
\href{https://eel.is/c++draft/dcl.spec.auto.general}{{[}dcl.spec.auto.general{]}}
& para 17 & Append an exception permitting \VERB|\KeywordTok{auto}| +
\VERB|\KeywordTok{override}|; point to
\href{https://eel.is/c++draft/class.virtual}{{[}class.virtual{]}} for
the resulting type. \\
2 & \href{https://eel.is/c++draft/class.virtual}{{[}class.virtual{]}} &
new para, after 8 & Fix the return type to the overridden function's;
make differing return types across multiple overridden functions
ill-formed. \\
3 & \href{https://eel.is/c++draft/cpp.predefined}{{[}cpp.predefined{]}}
& Table {[}tab:cpp.predefined.ft{]} & Add the
\VERB|\NormalTok{\_\_cpp\_deduced\_virtual\_override\_return}|
feature-test macro. \\
\end{longtable}
}

\section{\texorpdfstring{Impact on the
Standard\hyperref[impact-on-the-standard]{}}{Impact on the Standard}}\label{impact-on-the-standard}

This is a pure language extension. It:

\begin{itemize}
\tightlist
\item
  changes no existing valid program's meaning (code that does not
  combine \VERB|\KeywordTok{auto}| with \VERB|\KeywordTok{override}| is
  unaffected);
\item
  makes well-formed a category of declarations that are ill-formed today
  (\VERB|\KeywordTok{auto}\NormalTok{ f}\OperatorTok{()} \KeywordTok{override}|
  with no trailing-return-type on a virtual function);
\item
  requires no library changes;
\item
  requires no changes to the ABI or object model --- an
  \VERB|\KeywordTok{auto}\NormalTok{ … }\KeywordTok{override}| function
  has the same type, mangling, and vtable slot it would have had if its
  return type were written out by hand.
\end{itemize}

There are no breaking changes and no deprecations.

\section{\texorpdfstring{Implementation
experience\hyperref[implementation-experience]{}}{Implementation experience}}\label{implementation-experience}

The feature requires no novel analysis. A conforming compiler already:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  resolves which base function(s) a member overrides (to validate
  \VERB|\KeywordTok{override}| and covariance), and
\item
  has the overridden function's return type available at that point.
\end{enumerate}

Implementing the proposal amounts to: when a virtual function's declared
return type is the bare \VERB|\KeywordTok{auto}| placeholder and
\VERB|\KeywordTok{override}| is present, set the function's return type
to the overridden function's return type (after the existing override
resolution) instead of entering body-based deduction. No implementation
experience exists yet; the author believes a prototype in a major
compiler is straightforward and intends to pursue one.

\section{\texorpdfstring{Future
directions\hyperref[future]{}}{Future directions}}\label{future}

The following are intentionally \textbf{out of scope} for this paper but
are natural follow-ups:

\begin{itemize}
\tightlist
\item
  \textbf{\VERB|\KeywordTok{decltype}\OperatorTok{(}\KeywordTok{auto}\OperatorTok{)}|
  overrides.} Could be defined to copy the base type with its value
  category, though the practical benefit over \VERB|\KeywordTok{auto}|
  is small.
\item
  \textbf{Deduced covariant return types.} Allowing
  \VERB|\KeywordTok{auto}| to deduce a covariant type from the body
  (rather than copying the base type exactly) was considered and
  rejected for R0 in favor of a simpler rule; it could be revisited.
\item
  \textbf{Dropping the \VERB|\KeywordTok{override}| requirement.}
  Permitting deduction for any overriding function was rejected (see
  {[}design.trigger{]}); revisiting it would require a story for silent
  overrides and accidental coupling.
\end{itemize}

\section{\texorpdfstring{Acknowledgements\hyperref[acknowledgements]{}}{Acknowledgements}}\label{acknowledgements}

\emph{To be added.}

\section{\texorpdfstring{References\hyperref[references]{}}{References}}\label{references}

\section{\texorpdfstring{References\hyperref[bibliography]{}}{References}}\label{bibliography}

\protect\phantomsection\label{refs}
\begin{CSLReferences}{0}{0}
\bibitem[\citeproctext]{ref-N5046}
{[}N5046{]} 2026-05-12. Working Draft, Programming Languages --- C++.
\href{https://www.open-std.org/jtc1/sc22/wg21/docs/papers/2026/n5046.pdf}{
\CSLIndent{https://www.open-std.org/jtc1/sc22/wg21/docs/papers/2026/n5046.pdf}}

\end{CSLReferences}

\end{document}
